---
title: "assign1_kat_leigh_part1"
author: "Kat Leigh"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# load packages needed for analysis

remotes::install_github("coolbutuseless/ggpattern")
library(ggpattern)
library(tidyverse)
library(lubridate)
library(broom)
library(patchwork)
library(ggfortify)
library(readxl)
library(gghighlight)
library(plotly)
library(janitor)

```
Creating two graphs of mountain yellow-legged frog (Rana muscosa) counts using data from the Sierra Lakes Inventory Project contained in the Environmental Data Initiative repository.

About Data: This research endeavor ran from 1995-2002 and has supported research and management of Sierra Nevada aquatic ecosystems and their terrestrial interfaces. It describes the physical characteristics of and surveyed aquatic communities for > 8,000 lentic water bodies in the southern Sierra Nevada, including lakes, ponds, marshes, and meadows.

Data citation: Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3

```{r}
# read in data

raw_frogs<- read_excel("C:/Users/kll86/OneDrive/Documents/UCSB/esm244_assign1_kat_leigh/esm244_ass1/sierra_amphibians_esm244.xlsx") %>% 
  clean_names() %>% 
  mutate(across(where(is.character), tolower)) %>% # convert all characters to lowercase
  mutate(survey_date = ymd(survey_date)) %>% # recognize dates
  mutate(year = year(survey_date)) %>% # create a year column
  relocate(year) %>% # stick year column in front
  arrange(year) # order by year

```

Graph One: the total mountain yellow-legged frog (Rana muscosa) count each year across all water bodies, by life stage excluding the ‘EggMass’ level.

```{r}
# remove eggmass, group by year and lifestage

filter_frogs <- raw_frogs %>%
  filter(amphibian_life_stage != "EggMass",
         amphibian_species == "ramu")

summary_frogs <- filter_frogs %>%
  select(-amphibian_location, -lake_id) %>%  # remove unneeded columns
  group_by(year, amphibian_life_stage) %>% 
  summarise(count = sum(amphibian_number))

# make a plot

frog_plot <- ggplot(summary_frogs, aes(x= year, y= count))+
  geom_col(fill = "light blue")+
  facet_wrap(~amphibian_life_stage)+
  labs(title = str_wrap("Total counts across all lake locations per year for each lifestage", 30),
       x= "Year",
       y= "Count")+
  theme_minimal()+
  theme(plot.title = element_text(size=13))

frog_plot

```

Graph Two: the total counts (over all years of the study) of combined adult and subadult endangered mountain yellow-legged frogs (Rana muscosa) observed in the 5 lakes with the greatest total observed counts.

```{r}
# get the top lakes for ALL lifestages except Egg Mass

lakes <- filter_frogs %>%
  filter(amphibian_life_stage != "EggMass") %>% 
 group_by(lake_id) %>% 
  summarise(count =sum(amphibian_number)) %>% 
  arrange(-count) 

# top 5 counts = all lakes w/ count > 5726

top_lakes <- lakes %>% 
  filter(count > 5891)

# create data for graph

lake_graph_data <- filter_frogs %>% 
  filter(str_detect(amphibian_life_stage, "adult"),
         lake_id %in% top_lakes$lake_id) %>% 
  group_by(lake_id, amphibian_life_stage) %>% 
  summarise(count = sum(amphibian_number)) %>% 
  arrange(-count)

# find order of lakes

order_lakes <- filter_frogs %>% 
  filter(str_detect(amphibian_life_stage, "adult"),
         lake_id %in% top_lakes$lake_id) %>% 
  group_by(lake_id) %>% 
  summarise(count = sum(amphibian_number)) %>% 
  arrange(-count) %>%
  mutate(order = seq(1,5,1))

# apply order to graph data

lake_graph_data_ordered <- lake_graph_data %>% 
  mutate(orders = ifelse(lake_id == "50183", 1,
                         ifelse(lake_id == "10223", 2,
                                ifelse(lake_id == "10249", 3,
                                       ifelse(lake_id == "11209", 4,
                                              ifelse(lake_id == "10411", 5, 0)))))) %>% 
  #mutate(lake_id = paste("Lake", lake_id, sep=" ")) %>% removed since this makes the labels on the graph too long
  arrange(-orders)

  # graph with formatting to differentiate the lakes and the life stages

plot_lakes <- lake_graph_data_ordered %>% 
  ggplot(aes(x= reorder(lake_id, orders), y= count))+
  geom_col_pattern(aes(pattern_density = amphibian_life_stage, pattern_fill = as.factor(lake_id), pattern_color = as.factor(lake_id)),
                   pattern = 'stripe',
                   fill    = 'white',
                   color = 'black')+
  labs(title= str_wrap("Total counts of adults and subadults in lakes with highest frog abundance", 39),
       x= "Lake ID Number",
       y= "Count",
       pattern_density= "Life Stage",
       pattern_color= 'Lake ID Number')+
  theme_minimal()+
  theme(plot.title = element_text(size=13), legend.box = "horizontal")+
  guides(pattern_density = "legend", pattern_fill = "none", pattern_color = "none")

plot_lakes
  
```

Combine the two graphs into a single compound figure.

```{r}

# combine plots

combo_plots <- frog_plot | plot_lakes

# format the combo plot title size

test <- theme(plot.title= element_text(size= 18))

# create the final combo plot w/ appropriate title, subtitle, and caption sizes

combo_plot_final <- combo_plots + plot_annotation(
  title = str_wrap('Counts of mountain yellow-legged frogs (Rana muscosa) between 1995 and 2002', 60),
  caption = str_wrap('Top plot includes all lifestages except Egg Mass, and all lake locations. Lower plot contains only adult and subadult lifestages, and only counts for lakes where Rana muscosa are most abundant (i.e. where the total count at a lake exceeded 11 individuals). Lakes are given reference ID numbers in order to preserve confidentiality.', 130),
  theme= test)

combo_plot_final
  

```

